Benchmarking on Rigetti architecture

### 2022-11-07 txt/maths.tex
Update note with mathematical foundations

\newcommand{\eg}{{\emph{e.g.\/}}}
\newcommand{\ie}{{\emph{i.e.\/}}}
\newcommand{\ket}[1]{\ensuremath{|#1\rangle}}
\newcommand{\bra}[1]{\ensuremath{\langle#1|}}
\newcommand{\ketbra}[2]{\ensuremath{\ket{#1}\bra{#2}}}
\newcommand{\proj}[1]{\ensuremath{\ketbra{#1}{#1}}}
\newcommand{\braket}[2]{\ensuremath{\langle{#1}|{#2}\rangle}}
\newcommand{\floor}[1]{\ensuremath{\lfloor #1 \rfloor}}
\newcommand{\complexity}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\new}[1]{ \textcolor{red}{#1} }
\newcommand{\1}{{\rm 1\hspace{-0.9mm}l}}
\newcommand{\Id}{{\rm 1\hspace{-0.9mm}l}}
\newcommand{\connected}{\sim}
\newcommand{\SPAN}{\mathrm{span}}
\newcommand{\Lrm}{\ensuremath{\mathrm{L}}}
\newcommand{\Urm}{\ensuremath{\mathrm{U}}}
\newcommand{\ee}{\ensuremath{\mathrm{e}}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\ii}{\ensuremath{\mathrm{i}}}
\newcommand{\EE}{\mathcal{E}}
\newcommand{\XX}{\mathcal{X}}
\newcommand{\MM}{\mathcal{M}}
\newcommand{\NN}{\mathcal{N}}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\TT}{\mathcal{T}}
\newcommand{\PP}{\mathcal{P}}
\renewcommand{\SS}{\mathcal{S}}
\newcommand{\UU}{\mathcal{U}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\DU}{\mathcal{DU}}
\newcommand{\NOT}{\sigma_x}
\newcommand{\idop}[1][\XX]{\ensuremath{\1_{#1}}}
\newcommand{\diaguni}{\ensuremath{\mathcal{DU}}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{tr}}
%\DeclareMathOperator{\diag}{diag}
%\DeclareMathOperator{\diag}{diag}
\journal{SoftwareX}

@@

\subsubsection{Model specification}
The scenario of measurements discrimination assumes that we have access to an unknown measurement
device. One of the two measurements, either $\PP$
or $\mathcal{Q}$,  was used during an experiment. The only information we have that one of these measurements was 
performed. We want to determine an upper bound on the
probability of correct discrimination between $\PP$ and $\mathcal{Q}$. Moreover, we also want to
construct an optimal discrimination strategy to achieve the highest possible probability.

Discriminating between von Neumann measurements is valuable from the applications viewpoint. That is because, unlike other types of measurements, they can be implemented on current hardware.
So
here, we will consider the discrimination task between two von Neumann measurements, one of them on the computational basis, whereas the second one on basis 
$U$, where $U$ is a unitary matrix. 
Theoretical scenario was characterized in~\cite{puchala2018strategies}.
Such a scheme can be best described into the following steps 
and illustrated in
Fig.~\ref{fig:theoretical_scheme}.

\begin{figure}[h!]
	\centering
$  $	\includegraphics[scale=1.7]{pics/theoretical_scheme}
	\caption{Theoretical  scheme of discrimination  between von Neumann measurements $\PP_{U}$ and $\PP_\Id$. }
	\label{fig:theoretical_scheme}
\end{figure}

\begin{enumerate}
	\item We initialize the discriminator $\ket{\psi_{0}}$ on bipartite  system.
	\item We perform one of two known measurements (either $\PP_{\Id}$ or 
	$\PP_{U}$) on first part of the input state  $\ket{\psi_{0}}$, that means $\left(\PP \otimes \Id \right) (\proj{\psi_0})$.  
	\item We measure the  first system and obtain a label $i$. 
	\item We perform the conditional binary  (optimal) measurement $\PP_{V_i}$ on the second part based on the output label $i$. 
	\item  We measure the second  system of the input state  $\ket{\psi_{0}}$ and 
	make a decision based on the received label $j$. If $j=0$, then we decide 
	that the $\PP_U$ occurs. Otherwise, we decide that $\PP_{\Id}$ occurs. 
\end{enumerate}

>>>>

\newcommand{\eg}{{\emph{e.g.\/}}}
\newcommand{\ie}{{\emph{i.e.\/}}}
\newcommand{\ket}[1]{\ensuremath{|#1\rangle}}
\newcommand{\bra}[1]{\ensuremath{\langle#1|}}
\newcommand{\ketbra}[2]{\ensuremath{\ket{#1}\bra{#2}}}
\newcommand{\proj}[1]{\ensuremath{\ketbra{#1}{#1}}}
\newcommand{\braket}[2]{\ensuremath{\langle{#1}|{#2}\rangle}}
\newcommand{\floor}[1]{\ensuremath{\lfloor #1 \rfloor}}
\newcommand{\complexity}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\new}[1]{ \textcolor{red}{#1} }
\newcommand{\1}{{\rm 1\hspace{-0.9mm}l}}
\newcommand{\Id}{{\rm 1\hspace{-0.9mm}l}}
\newcommand{\connected}{\sim}
\newcommand{\SPAN}{\mathrm{span}}
\newcommand{\Lrm}{\ensuremath{\mathrm{L}}}
\newcommand{\Urm}{\ensuremath{\mathrm{U}}}
\newcommand{\ee}{\ensuremath{\mathrm{e}}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\ii}{\ensuremath{\mathrm{i}}}
\newcommand{\EE}{\mathcal{E}}
\newcommand{\XX}{\mathcal{X}}
\newcommand{\MM}{\mathcal{M}}
\newcommand{\NN}{\mathcal{N}}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\TT}{\mathcal{T}}
\newcommand{\PP}{\mathcal{P}}
\newcommand{\QQ}{\mathcal{Q}}
\renewcommand{\SS}{\mathcal{S}}
\newcommand{\UU}{\mathcal{U}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\DU}{\mathcal{DU}}
\newcommand{\NOT}{\sigma_x}
\newcommand{\idop}[1][\XX]{\ensuremath{\1_{#1}}}
\newcommand{\diaguni}{\ensuremath{\mathcal{DU}}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{tr}}
%\DeclareMathOperator{\diag}{diag}
%\DeclareMathOperator{\diag}{diag}
\journal{SoftwareX}

@@

\subsubsection{Motivation}

Suppose we have access to a device performing one of the predefined von Neumann
measurements\footnote{In PyQBench, we restrict ourselves to discriminating von Neumann measurements.
This is because, unlike other measurement types, they can be implemented on actual hardware.
},  $\PP$ or $\QQ$. While the $\PP$ and $\QQ$ are known, it is not known which of them is
performed when the device executes. Based on the measurement outcome, you have to guess which
measurement was performed (but you can perform arbitrary unitary operations before and after the
measurement). What is the highest probability of making a correct guess? And what do you have to do
to achieve it? And, most importantly, why would we want to do this?

Suppose that you know a strategy that, for an ideal device, would yield a probability
$p_{\text{succ}}$ of successfully discriminating between two measurements. Will the probability be
the same on an actual physical device? For current, Noisy Intermediate Scale Quantum devices
(NISQs), the answer is: certainly not. However, the error rate that you make when guessing can be
used as a benchmarking metric. PyQBench helps you in organizing such discrimination experiments for
a single qubit system\footnote{As we will soon see, the optimal discrimination strategy requires
the usage of an additional auxiliary qubit.}, executing them on real hardware or simulators, and
computing discrimination probabilities based on the measured bitstrings.

\subsubsection{Discrimination scheme}

Without loss of generality, we consider discrimination between a measurement in the computational
Z-basis ($\PP_\Id$), and an alternative measurement performed in the basis $U$
($\PP_U$)\footnote{Explaining why we can consider only Z-basis and alternative measurement is beyond
the scope of this technical documentation. See \cite{puchala2018strategies} if you are interested in
the explanation.}. In PyQBench, we operate only on two-level systems, but the discrimination scheme,
described in detail in \cite{puchala2018strategies}, makes no assumptions about the dimensionality.

In general, the discrimination scheme, presented in Fig.\ref{fig:theoretical_scheme}, requires a
second system of the same dimensionality as the measured one. First, the joint system is prepared in
some state $\ket{\psi_0}$. Then, the unknown measurement is performed on the first part of the
system. Based on its outcome $i$, another measurement $\mathcal{P}_{V_i}$ is performed to obtain an
outcome $j$. Finally, if $j=0$ we guess that the performed measurement is $\mathcal{P}_U$, otherwise
we guess that it was $\mathcal{P}_\Id$.

\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/theoretical_scheme}
	\caption{Theoretical  scheme of discrimination  between von Neumann measurements $\PP_{U}$ and $\PP_\Id$. }
	\label{fig:theoretical_scheme}
\end{figure}
